An Algorithm For Spatial Mapping Using a Hexagon Tile Map, With Application to Australian Maps

### 2020-01-18 kobakian_sugarbag_jss/kobakian_sugarbag_jss.Rmd
edits

The current practice for presenting geospatial data involves a choropleth map display. These maps highlight the geographic patterns in geospatially related statistics \citep{SAMGIS}.
The land on the map space is divided into geographic units, these boundaries are usually administrative, such as states or counties. The units are filled with colour to represent the value of the statistic \citep{EI}.

<!-- urbanisation -->
Australian residents are increasingly congregating around major cities, the vast rural areas are often sparsely populated in comparison to the urban centres. 
In Australia, government bodies such as the Australian Bureau of Statistics (ABS), and the Australian Electoral Commission (AEC) hold the responsibility for the division of geographic units. If it absolutely necessary, the AEC may adjust as the population increases. The division of the population into approximately equal population areas results in dramatically different square meterage of the geographic areas. This results in an unequal attention given to the statistic of each area, this can allow misrepresentation of the spatial distributions of human related statistics in geographic maps.

<!-- solutions -->
The solution to this visualisation problem begins with the geography. Cartograms apply a transformation to the geographic boundaries based on the value of the statistic of interest. These displays result in a distortion of the map space to represent differences in the statistic across the areas \citep{ACCAC}. 
The statistic of interest is used to determine the layout. When using the
Australian population, the result is a population cartogram that fails to preserve a recognisable display due to the difference in size of metropolitan and rural areas \citep{ACTUC}, \citep{GOINO}.
Contiguous cartograms change the shape of an area, preserving boundary relationships of neighbours. Non-contiguous cartograms maintain the geographic shape of each area, but lose the connection to neighbours as each areas shrinks or grows.

Alternative maps shift the focus from land area and shape, to the value of the statistics in a group of areas.
Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each area, or person is equally important.

Tilegrams, Rectangular cartograms \citep{ORC} and Dorling cartograms \citep{ACTUC}, all use one simple shape to represent each area. This allows preservation of spatial relationships and decreases the emphasis from the size of the geographic areas. These maps focus on the relationship between neighbours attempting to preserve connections, and disregard the unique shapes of the administrative boundaries.

The \pkg{sugarbag} package provides a new method to create tesselated hexagon tilegrams. 
Extending the tilegram to Australian applications required preserving the spatial relationships. It emphasises the capital cities as population hubs, and emphasises the distances rather than size of large, rural geographic units.

@@

This solution operates on a set of simple feature geometry objects, also known as \pkg{sf} \citep{sf} polygons.

There are four steps to create a tesselated hexagon tilegram.
These steps can be executed by the main function, \code{create_hexmap}, or can be implemented separately for more flexibility. There are parameters used in the process that can be provided, if they are not, they will be automatically derived.

@@

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2016} of Tasmania in 2016 is provided with the \pkg{sugarbag} package as \code{tas_sa2}. 
A single column of the data set must be used to identify the unique areas. 
In this case, the unique SA2 names for each SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are used to allocate areas around the closest capital city. Hobart will be the common focal point, as this example uses only the areas in the state of Tasmania.

@@

The following parameters will be determined within \code{create_hexmap} if they are not provided. They are created throughout the following example:

@@

Individual \pkg{sugarbag} functions can be used outside of the main function.
The set of polygons should be provided as an \pkg{sf} object, this is a data frame containing a \code{geometry} column. The \code{read_shape} function can assist in creating this object.

The centroids can be derived from the set of polygons using the \code{create_centroids} function:
```{r cents}
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
```

@@

A tilegram presents areas on a tellesated set of tiles.
The grid is created to ensure tessellation between the hexagons.

The grid of possible hexagon centroids is made using the \code{create_grid} function.
The grid creation requires several steps.
It uses the centroids, the hexagon size and the buffer distance.

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is defined by the size specified as \code{hex_size}.
The minimum and maximum, longitude and latitude values of the centroid locations are found. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. The rows are then created from the latitude values and the buffer distance.
An individual hexagon location is created from all intersections of the longitude columns and latitude rows.

A square grid could be used for sqaure tiles, but it will not facilitate tessellated hexagons. Figure \ref{fig:grid2} allows for hexagons, as every second latitude row on the grid is shifted right, by half of the hexagon size.

```{r grid2, fig.cap = "Grid points to create a tilegram.", out.width = "100%", echo = FALSE}
knitr::include_graphics(path = "figures/2grid.png")
```

@@

Not all of the grid points will be used, especially if islands result in a large grid space.
To filter the grid for appropriate points for allocation,
the \code{create_buffer} function is called within \code{create_grid}.
It finds the grid points needed to best capture the set of centroids on a hexagon tile map.

For each centroid location, the closest latitude row and longitude column are found. Then rows and columns of centroids are divided into 20 groups.
The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The rolling windows can be seen on the bottom and right of Fig. \ref{fig:filter_grid}.
This will tailor the available grid points to those most likely to be used.
It also helps reduce the amount of time taken, as it decreases the amount of points considered for each centroid allocation.

@@

Only the grid points between the rolling average of the minimum and maximum centroid values are kept, for each row and column of the grid.

```{r filter_grid, fig.cap = "All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots shown the grid points left as to choose from after the buffer step. The rolling windows to the right show the rows used to filter the hexagon locations.", echo = FALSE, out.width = "100%"}
knitr::include_graphics(path = "figures/3grid.png", dpi = 300)
```

@@

The distance between each centroid in the set, and each of the focal points provided is calculated. The name of the closest focal point, and the distance and angle from focal point to polygon centroid is joined to polygon data set. To minimise time taken for this step, only Tasmania's capital city Hobart is provided.
The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. The points are arranged from the centroid closest to the focal points, to the furthest.

@@

Allocation of all centroids takes place using the set of polygon centroids and the hexagon map grid. 
Centroid allocation begins with the closest centroid to a focal point.
This will preserve spatial relationships with the focal point, as the inner city areas are allocated first, they will be placed closest to the capital, and the areas that are further will then be accommodated.
Only the hexagon grid points that have not yet been allocated are considered.

The possible hexagon locations consider for a centroid location are determined by the \code{hex_filter}. This is the maximum amount of hexagons between the centroid and the furthest considered hexagon. It is used to subset possible grid points to only those surrounding the polygon centroid within an appropriate range. A smaller distance will increase speed, but can decrease accuracy if the angle width increases.

@@

The following example considers one of the Statistical Areas at Level 2. 
Within the algorithm, these steps are repeated for each polygon.

@@

Keep only the available hexagon points, this will prevent multiple areas being allocated to the same hexagon.

@@

This will allow only the closest points that are not yet assigned, to be considered.

A box of possible hexagon locations around the centroid. The corners of the box may not look square as the buffer has already removed unnecessary points from over the ocean.

@@

The \code{width} parameter is used to take a slice of the remaining points. This uses the angle from the closest capital city, to the current centroid. This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

If no available hexagon grid point is found within the original filter distance and angle, the distance is expanded, only when a maximum distance is reached will the angle expand to accommodate more possible grid points.  
By default the angle filter to hexagon grid points that fall within the bounds of the angle from the focal point to the geographic centroid, plus and minus 30 degrees. This will increase if no points can be found within the \code{hex_filter} distance. The angle of 30 was chosen to allow the algorithm to choose hexagons that best maintained the spatial relationship between the focal point and geographic centroid.
The allocation is returned and combined with the data relating to each polygon.

@@

A complete hexagon tile map of Tasmania is created by applying the algorithm steps to each centroid.
The hexagon tile map visualisation is used below to visualise the Australian Cancer Atlas data. Two views of the same data are produced by filling according to the Lung Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in Figure  \ref{fig:sir} shows the group of blue areas in the Hobart CBD more prominently in the hexagon tile map (b).
The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the more rural areas.

@@

It is possible to incorporate the neighbouring areas for each SA2, for stronger preservation of the spatial distribution.

An additional step can be included to allow the neighbours that have alredy been allocated to influence the placement of the current centroid. 
This requires respecifying the \code{sf} object as the argument for the \code{use_neighbours} parameter. This calculates neighbours using intersections of their polygons. This occurs for all areas before any allocations begin.

For the current centroid, the list of neighbours is consulted. If any neighbour was already allocated, the surrounding hexagons on the grid are prioritised. For multiple neighbours, the neighbouring hexagon grid points are aggregated and considered in order of distance from the original centroid.

@@

The \pkg{gganimate} package can be used to make an animation.
It requires connecting the polygons for each area in two displays two displays, which can be done using the \code{sf_id} variable, such as the SA2 name.
The animation^[This animation can be viewed at: https://sugarbagjss.netlify.com/] connecting these two displays will highlight the rapid growth of the inner-city areas, and will decrease the large rural areas. The hexagons that move the furthest will move rapidly in the animation.

@@

It is possible to use alternative maps to communicate spatial distributions.While a choropleth map display is the current practice spatial visualisation of geographical data. Current methods do not always work for Australia due to the large gaps between densely populated capital cities. The administrative boundaries may distract from the statistics, communicated using colour.

Alternative maps highlight the value of the statistics across the geographic units. Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem has equally sized areas, with neighbourhood boundary connections. This map algorithm is implemented in the \pkg{sugarbag} @sugarbag package written for \code{R} @R.
The \pkg{sugarbag} package creates tesselated hexagon tilegrams. The Australian application preserves the spatial relationships, emphasising capital cities. The hexagon tile map is a visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tesselation does not allow the size of the hexagons to represent another variable, similar to the choropleth maps.
The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. It works on the assumption that viewers can use directional assumptions to identify their neighbourhoods, this can be aided by the animation.

@@

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualizations. They would also like to thank Professor Kerrie Mengersen for suggestions and comments. Mitch O'Hara-Wild provided assistance with some parts of the algorithm and code. Sayani Gupta provided helpful advice on parts of this paper.

@@

Australian Electoral Data can be found in the \pkg{eechidna} package for R.

>>>>

The current practice for presenting geospatial data is a choropleth map display. These maps highlight the geographic patterns in geospatially related statistics \citep{SAMGIS}.
The land on the map space is divided into geographic units, these boundaries are usually administrative, such as states or counties. The units are filled with colour to represent the value of the statistic \citep{EI}.

<!-- urbanisation -->
Australian residents are increasingly congregating around major cities, the vast rural areas are often sparsely populated in comparison to the urban centres. 
In Australia, government bodies such as the Australian Bureau of Statistics (ABS), and the Australian Electoral Commission (AEC) hold the responsibility for the division of the population into geographic units. If it necessary, the AEC may adjust the boundaries of the areas as the population increases. The division of the population into approximately equal population areas results in dramatically different square meterage of the geographic areas. This can give unequal attention to the statistic of each area, this can cause misrepresentation of the spatial distributions of human related statistics in geographic maps.

<!-- solutions -->
The solutions to this visualisation problem begin with the geography. Cartograms apply a transformation to the geographic boundaries based on the value of the statistic of interest. These displays result in a distortion of the map space to represent differences in the statistic across the areas \citep{ACCAC}. 
The statistic of interest is used to determine the cartogram layout. When the
Australian population is the statistic of interest, the result is a population cartogram. They fail to preserve a recognisable display due to the difference in size of metropolitan and rural areas \citep{ACTUC}, \citep{GOINO}.
Contiguous cartograms change the shape of areas, while preserving boundary relationships of neighbours. Non-contiguous cartograms maintain the geographic shape of each geographic area, but will lose the connection to neighbours as the polygon for each geographic area shrinks or grows.

Alternative maps shift the focus from land area and shape, to the value of the statistics in a group of areas.
Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each area, or person within it is equally important.

tile maps, Rectangular cartograms \citep{ORC} and Dorling cartograms \citep{ACTUC}, all use one simple shape to represent each area. They place various importance on the preservation of spatial relationships, but all decrease the emphasis on the size of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries.

The \pkg{sugarbag} package provides a new algorithm to create tesselated hexagon tile maps. 
It emphasises the capital cities as population hubs, and emphasises the distances rather than size of large, rural geographic units.

@@

The algorithm presented in \pkg{sugarbag} package operates on a set of simple feature geometry objects, also known as \pkg{sf} \citep{sf} polygons.

There are four steps performed to create a tesselated hexagon tile map.
These steps can be executed by the main function, \code{create_hexmap}, or can be implemented separately for more flexibility. There are parameters used in the process that can be provided by users, if they are not, they will be automatically derived.

@@

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2016} of Tasmania in 2016 is provided with the \pkg{sugarbag} package as \code{tas_sa2}. 
A single column of the data set is used to identify the unique areas. 
In this case, the unique SA2 names for each SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are used as focal points to allocate each geographic area around the closest capital city. Hobart will be the common focal point, as this example uses only the areas in the state of Tasmania.

@@

The following parameters will be determined within \code{create_hexmap} if they are not provided. They are created as they are needed throughout the following example:

@@

A set of centroids may be used directly. The set of polygons should be provided as an \pkg{sf} object, this is a data frame containing a \code{geometry} column. The \code{read_shape} function can assist in creating this object for use in `R`.

The centroids can be derived from the set of polygons using the \code{create_centroids} function:

```{r cents}
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
```

@@

A grid is created to ensure tessellation between the hexagons that represent the geographic units on a hexagon tile map.

The grid of possible hexagon locations is made using the \code{create_grid} function.
It uses the centroids, the hexagon size and the buffer distance.

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by \code{hex_size}.
Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance.
A unique hexagon location is created from all intersections of the longitude columns and latitude rows.
Figure \ref{fig:grid2} shows the original grid on the left, to allow for tessellating hexagons, every second latitude row on the grid is shifted right, by half of the hexagon size. The grid for tessellation is shown on the right.

```{r grid2, fig.cap = "Grid points to create a tile map.", out.width = "100%", echo = FALSE}
knitr::include_graphics(path = "figures/2grid.png")
```

@@

Not all of the grid points will be used, especially if islands result in a large grid space.
To filter the grid for appropriate hexagon locations for allocation, the \code{create_buffer} function is used by \code{create_grid}.
It finds the grid points needed to best capture the set of centroids on a hexagon tile map.

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups.
The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The rolling windows can be seen on the bottom and right of the gris shown in Fig. \ref{fig:filter_grid}.
This will tailor the available grid points to those most likely to be used.
It also helps reduce the amount of time taken, as it decreases the amount of points considered for each centroid allocation.

@@

The grid points are kept only if they fall between the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. The sparsely populated South-West region of National Park has much fewer points available compared to the South-East region containing the city of Hobart.

```{r filter_grid, fig.cap = "All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots show the grid points left to choose from after the buffer step. The rolling windows show the collections of rows and columns used to filter the hexagon locations.", echo = FALSE, out.width = "100%"}
knitr::include_graphics(path = "figures/3grid.png", dpi = 300)
```

@@

The distance between each centroid in the set, and each of the focal points provided is calculated. The name of the closest focal point, and the distance and angle from focal point to polygon centroid is joined to polygon data set. To minimise time taken for this step only one option is provided, Tasmania's capital city Hobart.
The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. The points are arranged from the centroid closest to the focal point(s), to the furthest.

@@

Allocation of all centroids takes place using the set of polygon centroids and the hexagon map grid. 
Centroid allocation begins with the closest centroid to a focal point.
This will preserve spatial relationships with the focal point, as the inner city areas are allocated first, they will be placed closest to the capital, and the areas that are further will then be accommodated.
The possible hexagon grid points reduces by one after each allocation, then only those that have not yet been allocated are considered.

The possible hexagon locations to consider for a centroid are determined by the \code{hex_filter}. This is the maximum amount of hexagons between the centroid and the furthest considered hexagon. It is used to subset possible grid points to only those surrounding the polygon centroid within an appropriate range. A smaller distance will increase speed, but can decrease accuracy when width of the angle increases.

@@

The following example considers the first of the Statistical Areas at Level 2. 
Within the algorithm, these steps are repeated for each polygon.

@@

Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon.

@@

A box of possible hexagon locations around the centroid allows only the closest points that are not yet assigned to be considered. The corners of the box may not appear square if the buffer step has already removed unnecessary points from over the ocean.

@@

The \code{width} parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \ref{fig:angles} . This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

```{r angles, fig.cap = "Filter for grid points within the angle from the focal point to the centroid.", echo = FALSE, out.width = "100%"}
knitr::include_graphics(path = "figures/5allocate.png")
```

If no available hexagon grid point is found within the original filter distance and angle, the distance is expanded, only when a maximum distance is reached will the angle expand to accommodate more possible grid points.  
By default the angle filter to hexagon grid points that fall within the bounds of the angle from the focal point to the geographic centroid, plus and minus 30 degrees. This will increase if no points can be found within the \code{hex_filter} distance. The default angle of 30 was chosen to allow the algorithm to choose hexagons that best maintained the spatial relationship between the focal point and geographic centroid.

@@

A complete hexagon tile map of Tasmania is created by applying the algorithm steps to each centroid.
The hexagon tile map visualisation is used below to visualise the Australian Cancer Atlas data. Two views of the same data are produced by filling according to the Lung Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in Figure \ref{fig:sir} shows the group of blue areas in the Hobart CBD more prominently in the hexagon tile map (b).
The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the rural areas.

@@

It is possible to consider the neighbouring areas for each SA2, for stronger preservation of the spatial distribution.

An additional step can be included to allow the neighbours that have alredy been allocated to influence the placement of the current centroid. 
This requires specifying the \code{sf} object as the argument for the \code{use_neighbours} parameter. This calculates neighbours using intersections of their polygons. This occurs for all areas before any allocations begin.

During the allocation of each centroid, the list of neighbours is consulted. If any neighbour was already allocated, the hexagons surrounding the neighbours on the grid are prioritised. For multiple neighbours, the neighbouring hexagon grid points are aggregated and considered in order of distance from the original centroid.

@@

The \pkg{gganimate} package can be used to make an animation.
It requires connecting the polygons for each area in two displays, which can be done using the \code{sf_id} variable, such as the SA2 name.
The animation^[This animation can be viewed at: https://sugarbagjss.netlify.com/] connecting these two displays will highlight the rapid growth of the inner-city areas, and will decrease the large rural areas. The hexagons that move the furthest will move rapidly in the animation.

@@

It is possible to use alternative maps to communicate spatial distributions.While a choropleth map display is the current practice spatial visualisation of geographical data. Current methods do not always work for Australia due to the large geographic space between the densely populated capital cities. The administrative boundaries may also distract from the statistics communicated using colour.

Alternative maps highlight the value of the statistics across the geographic units. Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem has equally sized areas, with neighbourhood boundary connections. This map algorithm is implemented in the \pkg{sugarbag} @sugarbag package written for \code{R} @R.
The \pkg{sugarbag} package creates tesselated hexagon tile maps. The Australian application preserves the spatial relationships, emphasising capital cities. The hexagon tile map is a visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tesselation does not allow the size of the hexagons to represent another variable, similar to the choropleth maps.
The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. It works on the assumption that viewers can use directional relationships to identify their neighbourhoods but this can be aided by the animation.

@@

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualizations. They would also like to thank Professor Kerrie Mengersen for suggestions and comments. Mitchell O'Hara-Wild provided assistance with some parts of the algorithm and code. Sayani Gupta provided helpful advice on parts of this paper.

@@

### 2020-01-18 kobakian_sugarbag_jss/kobakian_sugarbag_jss.tex
pdf files
edits

\Plainauthor{Stephanie Kobakian, Dianne Cook, Earl Duncan}
\Plaintitle{An Algorithm For Spatial Mapping Using a Hexagon Tile Map, With
Application to Australian Maps}

@@

@@

The current practice for presenting geospatial data involves a
choropleth map display. These maps highlight the geographic patterns in
geospatially related statistics \citep{SAMGIS}. The land on the map
space is divided into geographic units, these boundaries are usually
administrative, such as states or counties. The units are filled with
colour to represent the value of the statistic \citep{EI}.

Australian residents are increasingly congregating around major cities,
the vast rural areas are often sparsely populated in comparison to the
urban centres. In Australia, government bodies such as the Australian
Bureau of Statistics (ABS), and the Australian Electoral Commission
(AEC) hold the responsibility for the division of geographic units. If
it absolutely necessary, the AEC may adjust as the population increases.
The division of the population into approximately equal population areas
results in dramatically different square meterage of the geographic
areas. This results in an unequal attention given to the statistic of
each area, this can allow misrepresentation of the spatial distributions
of human related statistics in geographic maps.

The solution to this visualisation problem begins with the geography.
Cartograms apply a transformation to the geographic boundaries based on
the value of the statistic of interest. These displays result in a
distortion of the map space to represent differences in the statistic
across the areas \citep{ACCAC}. The statistic of interest is used to
determine the layout. When using the Australian population, the result
is a population cartogram that fails to preserve a recognisable display
due to the difference in size of metropolitan and rural areas
\citep{ACTUC}, \citep{GOINO}. Contiguous cartograms change the shape of
an area, preserving boundary relationships of neighbours. Non-contiguous
cartograms maintain the geographic shape of each area, but lose the
connection to neighbours as each areas shrinks or grows.

Alternative maps shift the focus from land area and shape, to the value
of the statistics in a group of areas. Alternative mapping methods allow
increased understanding of the spatial distribution of a variable across
the population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each area, or person is equally important.

Tilegrams, Rectangular cartograms \citep{ORC} and Dorling cartograms
\citep{ACTUC}, all use one simple shape to represent each area. This
allows preservation of spatial relationships and decreases the emphasis
from the size of the geographic areas. These maps focus on the
relationship between neighbours attempting to preserve connections, and
disregard the unique shapes of the administrative boundaries.

The \pkg{sugarbag} package provides a new method to create tesselated
hexagon tilegrams. Extending the tilegram to Australian applications
required preserving the spatial relationships. It emphasises the capital
cities as population hubs, and emphasises the distances rather than size
of large, rural geographic units.

@@

Individual \pkg{sugarbag} functions can be used outside of the main
function. The set of polygons should be provided as an \pkg{sf} object,
this is a data frame containing a \code{geometry} column. The
\code{read_shape} function can assist in creating this object.

@@

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
defined by the size specified as \code{hex_size}. The minimum and
maximum, longitude and latitude values of the centroid locations are
found. Equally spaced columns are created from the minimum longitude
minus the buffer distance, up to the maximum longitude plus the buffer
distance. The rows are then created from the latitude values and the
buffer distance. An individual hexagon location is created from all
intersections of the longitude columns and latitude rows.

A square grid could be used for sqaure tiles, but it will not facilitate
tessellated hexagons. Figure \ref{fig:grid2} allows for hexagons, as
every second latitude row on the grid is shifted right, by half of the
hexagon size.

@@

\caption[Grid points to create a tilegram]{Grid points to create a tilegram.}\label{fig:grid2}
\end{figure}
\end{CodeChunk}

@@

Only the grid points between the rolling average of the minimum and
maximum centroid values are kept, for each row and column of the grid.

@@

\caption[All possible hexagon locations from the initial grid are shown with blue outlines]{All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots shown the grid points left as to choose from after the buffer step. The rolling windows to the right show the rows used to filter the hexagon locations.}\label{fig:filter_grid}
\end{figure}
\end{CodeChunk}

@@

The \code{width} parameter is used to take a slice of the remaining
points. This uses the angle from the closest capital city, to the
current centroid. This allows the spatial relationship to be preserved,
even when it is allocated to a hexagon that is further from the focal
point then the original centroid location.

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded, only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points.\\
By default the angle filter to hexagon grid points that fall within the
bounds of the angle from the focal point to the geographic centroid,
plus and minus 30 degrees. This will increase if no points can be found
within the \code{hex_filter} distance. The angle of 30 was chosen to
allow the algorithm to choose hexagons that best maintained the spatial
relationship between the focal point and geographic centroid. The
allocation is returned and combined with the data relating to each
polygon.

@@

The \pkg{gganimate} package can be used to make an animation. It
requires connecting the polygons for each area in two displays two
displays, which can be done using the \code{sf_id} variable, such as the
SA2 name. The animation\footnote{This animation can be viewed at:
  \url{https://sugarbagjss.netlify.com/}} connecting these two displays
will highlight the rapid growth of the inner-city areas, and will
decrease the large rural areas. The hexagons that move the furthest will
move rapidly in the animation.

@@

It is possible to use alternative maps to communicate spatial
distributions.While a choropleth map display is the current practice
spatial visualisation of geographical data. Current methods do not
always work for Australia due to the large gaps between densely
populated capital cities. The administrative boundaries may distract
from the statistics, communicated using colour.

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \pkg{sugarbag} \citet{sugarbag} package written for \code{R}
\citet{R}. The \pkg{sugarbag} package creates tesselated hexagon
tilegrams. The Australian application preserves the spatial
relationships, emphasising capital cities. The hexagon tile map is a
visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tesselation
does not allow the size of the hexagons to represent another variable,
similar to the choropleth maps. The algorithm is heavily dependent on
the focal points used, as this determines the order of allocation. It
works on the assumption that viewers can use directional assumptions to
identify their neighbourhoods, this can be aided by the animation.

>>>>

\Plainauthor{Stephanie Kobakian, Dianne Cook, Earl Duncan}
\Plaintitle{An Algorithm For Spatial Mapping Using a Hexagon Tile Map, With
Application to Australian Maps}
\Shorttitle{An Algorithm For Spatial Mapping Using a Hexagon Tile Map}

@@

\setcounter{page}{65}

@@

The current practice for presenting geospatial data is a choropleth map
display. These maps highlight the geographic patterns in geospatially
related statistics \citep{SAMGIS}. The land on the map space is divided
into geographic units, these boundaries are usually administrative, such
as states or counties. The units are filled with colour to represent the
value of the statistic \citep{EI}.

Australian residents are increasingly congregating around major cities,
the vast rural areas are often sparsely populated in comparison to the
urban centres. In Australia, government bodies such as the Australian
Bureau of Statistics (ABS), and the Australian Electoral Commission
(AEC) hold the responsibility for the division of the population into
geographic units. If it necessary, the AEC may adjust the boundaries of
the areas as the population increases. The division of the population
into approximately equal population areas results in dramatically
different square meterage of the geographic areas. This can give unequal
attention to the statistic of each area, this can cause
misrepresentation of the spatial distributions of human related
statistics in geographic maps.

The solutions to this visualisation problem begin with the geography.
Cartograms apply a transformation to the geographic boundaries based on
the value of the statistic of interest. These displays result in a
distortion of the map space to represent differences in the statistic
across the areas \citep{ACCAC}. The statistic of interest is used to
determine the cartogram layout. When the Australian population is the
statistic of interest, the result is a population cartogram. They fail
to preserve a recognisable display due to the difference in size of
metropolitan and rural areas \citep{ACTUC}, \citep{GOINO}. Contiguous
cartograms change the shape of areas, while preserving boundary
relationships of neighbours. Non-contiguous cartograms maintain the
geographic shape of each geographic area, but will lose the connection
to neighbours as the polygon for each geographic area shrinks or grows.

Alternative maps shift the focus from land area and shape, to the value
of the statistics in a group of areas. Alternative mapping methods allow
increased understanding of the spatial distribution of a variable across
the population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each area, or person within it is equally important.

tile maps, Rectangular cartograms \citep{ORC} and Dorling cartograms
\citep{ACTUC}, all use one simple shape to represent each area. They
place various importance on the preservation of spatial relationships,
but all decrease the emphasis on the size of the geographic areas. These
alternative map displays focus on the relationship between neighbours,
attempting to preserve connections, and disregard the unique shapes of
the administrative boundaries.

The \pkg{sugarbag} package provides a new algorithm to create tesselated
hexagon tile maps. It emphasises the capital cities as population hubs,
and emphasises the distances rather than size of large, rural geographic
units.

@@

A set of centroids may be used directly. The set of polygons should be
provided as an \pkg{sf} object, this is a data frame containing a
\code{geometry} column. The \code{read_shape} function can assist in
creating this object for use in \texttt{R}.

@@

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \code{hex_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:grid2} shows the original
grid on the left, to allow for tessellating hexagons, every second
latitude row on the grid is shifted right, by half of the hexagon size.
The grid for tessellation is shown on the right.

@@

\caption[Grid points to create a tile map]{Grid points to create a tile map.}\label{fig:grid2}
\end{figure}
\end{CodeChunk}

@@

The grid points are kept only if they fall between the rolling average
of the minimum and maximum centroid values after accounting for the
buffer distance, for each row and column of the grid. The sparsely
populated South-West region of National Park has much fewer points
available compared to the South-East region containing the city of
Hobart.

@@

\caption[All possible hexagon locations from the initial grid are shown with blue outlines]{All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots show the grid points left to choose from after the buffer step. The rolling windows show the collections of rows and columns used to filter the hexagon locations.}\label{fig:filter_grid}
\end{figure}
\end{CodeChunk}

@@

The \code{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:angles} . This allows the spatial relationship to be preserved,
even when it is allocated to a hexagon that is further from the focal
point then the original centroid location.

{\centering \includegraphics[width=1\linewidth]{figures/5allocate}

\caption[Filter for grid points within the angle from the focal point to the centroid]{Filter for grid points within the angle from the focal point to the centroid.}\label{fig:angles}
\end{figure}
\end{CodeChunk}

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded, only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points.\\
By default the angle filter to hexagon grid points that fall within the
bounds of the angle from the focal point to the geographic centroid,
plus and minus 30 degrees. This will increase if no points can be found
within the \code{hex_filter} distance. The default angle of 30 was
chosen to allow the algorithm to choose hexagons that best maintained
the spatial relationship between the focal point and geographic
centroid.

@@

The \pkg{gganimate} package can be used to make an animation. It
requires connecting the polygons for each area in two displays, which
can be done using the \code{sf_id} variable, such as the SA2 name. The
animation\footnote{This animation can be viewed at:
  \url{https://sugarbagjss.netlify.com/}} connecting these two displays
will highlight the rapid growth of the inner-city areas, and will
decrease the large rural areas. The hexagons that move the furthest will
move rapidly in the animation.

@@

It is possible to use alternative maps to communicate spatial
distributions.While a choropleth map display is the current practice
spatial visualisation of geographical data. Current methods do not
always work for Australia due to the large geographic space between the
densely populated capital cities. The administrative boundaries may also
distract from the statistics communicated using colour.

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \pkg{sugarbag} \citet{sugarbag} package written for \code{R}
\citet{R}. The \pkg{sugarbag} package creates tesselated hexagon tile
maps. The Australian application preserves the spatial relationships,
emphasising capital cities. The hexagon tile map is a visualisation
solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tesselation
does not allow the size of the hexagons to represent another variable,
similar to the choropleth maps. The algorithm is heavily dependent on
the focal points used, as this determines the order of allocation. It
works on the assumption that viewers can use directional relationships
to identify their neighbourhoods but this can be aided by the animation.
